---
title: "PAC report"
output: html_document
name: "Charlotte Lin"
date: "2023-12-07"
---


#Exploring the data (EDA) / Preparing the data (Data cleaning)

```{r}
getwd()
setwd("/Users/tinalin/Downloads")
data = read.csv('analysisData.csv')

library(xgboost)
library(vtreat)
library(ggplot2)
library(dplyr)
library(corrplot)
library(broom)
library(caret)
library(car)
library(leaps)
library(stats)
library(gbm)

str(data)

#first attempt - only select numerical variables
d<- data %>%
  select_if(~is.numeric(.))%>%
  mutate(total_volume=length_inches*width_inches*height_inches)

cdata<- data %>%
  mutate(total_volume=length_inches*width_inches*height_inches)%>%
  mutate(torque_n=substr(torque, 1, 3))

#make categorical to numerical
cdata$is_new <- as.integer(as.logical(cdata$is_new))
cdata$make_name <- as.numeric(as.factor(cdata$make_name))
cdata$listing_color <- as.numeric(as.factor(cdata$listing_color))
cdata$torque_n <- as.integer(cdata$torque_n)

#replace NAs with median
cdata$owner_count[is.na(cdata$owner_count)] <- median(cdata$owner_count, na.rm = TRUE)
cdata$mileage[is.na(cdata$mileage)] <- median(cdata$mileage, na.rm = TRUE)
cdata$engine_displacement[is.na(cdata$engine_displacement)] <- median(cdata$engine_displacement, na.rm = TRUE)
cdata$highway_fuel_economy[is.na(cdata$highway_fuel_economy)] <- median(cdata$highway_fuel_economy, na.rm = TRUE)
cdata$horsepower[is.na(cdata$horsepower)] <- median(cdata$horsepower, na.rm = TRUE)
cdata$fuel_tank_volume_gallons[is.na(cdata$fuel_tank_volume_gallons)] <- median(cdata$fuel_tank_volume_gallons, na.rm = TRUE)
cdata$seller_rating[is.na(cdata$seller_rating)] <- median(cdata$seller_rating, na.rm = TRUE)
cdata$city_fuel_economy[is.na(cdata$city_fuel_economy)] <- median(cdata$city_fuel_economy, na.rm = TRUE)
cdata$torque_n[is.na(cdata$torque_n)] <- median(cdata$torque_n, na.rm = TRUE)

car_data <- d[complete.cases(d), ]
cor(car_data)

#corrplot(d_,method = "color",order="alphabet")
#corrplot(d_cor,order="AOE")
#corrplot(d_cor,method = "number",number.cex = 0.5)
#corrplot(carp,method="number")

carr<- cdata%>%
  select_if(~is.numeric(.))%>%
  subset(select=-c(length_inches,width_inches,height_inches))

```

 #Analysis technique (Feature selection, Model Selection)

```{r}

set.seed(1031)
split = createDataPartition(y=carr$price,p = 0.8,list = F,groups = 100)
train = carr[split,]
test = carr[-split,]

subsets = regsubsets(price~.,data=train, nvmax=21)
summary(subsets)

subsets_measures = data.frame(model=1:length(summary(subsets)$cp),
                              cp=summary(subsets)$cp,
                              bic=summary(subsets)$bic, 
                              adjr2=summary(subsets)$adjr2)
subsets_measures
#based on CP - 21 variables
#based on BIC - 17 variables
#based on adjr2 - 20 variables

start_mod = lm(price~.,data=train)
empty_mod = lm(price~1,data=train)
full_mod = lm(price~.,data=train)
backwardStepwise = step(start_mod,
                        scope=list(upper=full_mod,lower=empty_mod),
                        direction='backward')


#XGBoost

features = c('year','horsepower','mileage','owner_count','listing_color','maximum_seating','make_name','is_new','wheelbase_inches','back_legroom_inches','front_legroom_inches','engine_displacement','total_volume','daysonmarket','seller_rating','torque_n','fuel_tank_volume_gallons','highway_fuel_economy')
target = 'price'

xgtrain = xgb.DMatrix(data = as.matrix(train[, features]),
                      label = train[, target])
xgtest = xgb.DMatrix(data = as.matrix(test[, features]),
                     label = test[, target])

model = xgboost(data=xgtrain, nrounds=2000, verbose = 1)

```

#Other models I've used

I also used Boosting with XGBoost, multiple regression and tuned random forest. (Tuned random forest did not run at all, so I already deleted the code) Besides, I spent a lot of time trying our different variable combinations and avoiding overfitting.

```{r}

##Boosting with XGBoost
trt = designTreatmentsZ(dframe = train,
                        varlist = names(train)[c(2:18,20,21)])
newvars = trt$scoreFrame[trt$scoreFrame$code%in% c('clean','lev'),'varName']

#prepare data for XGBoost
train_input = prepare(treatmentplan = trt, 
                      dframe = train,
                      varRestriction = newvars)
test_input = prepare(treatmentplan = trt, 
                     dframe = test,
                     varRestriction = newvars)

head(train_input)

#Tune XGBoost Parameters
set.seed(1031)
tune_nrounds = xgb.cv(data=as.matrix(train_input), 
                      label = train$price,
                      nrounds=250,
                      nfold = 5,
                      verbose = 0)

#Visualize Tuning Results
ggplot(data=tune_nrounds$evaluation_log, aes(x=iter, y=test_rmse_mean))+
  geom_point(size=0.4, color='sienna')+
  geom_line(size=0.1, alpha=0.1)+
  theme_bw()

#Identify Optimal Number of Rounds
which.min(tune_nrounds$evaluation_log$test_rmse_mean)

#Train XGBoost Model
xgboost2= xgboost(data=as.matrix(train_input), 
                  label = train$price,
                  nrounds=249,
                  verbose = 0)

#Evaluate XGBoost Model on Test Data
pred = predict(xgboost2, 
               newdata=as.matrix(test_input))

rmse_xgboost = sqrt(mean((pred - test$price)^2))
rmse_xgboost ## 3420.524


## Multiple Regression
model1=lm(price~fuel_tank_volume_gallons + highway_fuel_economy + city_fuel_economy + 
            front_legroom_inches + total_volume + 
            engine_displacement + horsepower + maximum_seating + year + 
            mileage + owner_count + seller_rating,data=train)

pred=predict(model1)
sse = sum((pred - train$price)^2)
sst = sum((mean(train$price)-train$price)^2)
model1_r2 = 1 - sse/sst; model1_r2

rmse = sqrt(mean((pred-train$price)^2)); rmse #rsme=7504.565 (model1 train)

predt = predict(model1, newdata=test)
sse_test = sum((predt - test$price)^2)
sst_test = sum((mean(train$price)-test$price)^2)
model1_r2_test = 1 - sse_test/sst_test; 
rmse_test = sqrt(mean((predt-test$price)^2));

rmse_test #rsme_test=7817.461 (model1 test)

```

```{r}

getwd()
setwd("/Users/tinalin/Desktop")
scoringData = read.csv("scoringData.csv")

scoringData1<- scoringData %>%
  mutate(total_volume=length_inches*width_inches*height_inches)%>%
  mutate(torque_n=substr(torque, 1, 3))

#make categorical to numerical
scoringData1$is_new <- as.integer(as.logical(scoringData1$is_new))
scoringData1$make_name <- as.numeric(as.factor(scoringData1$make_name))
scoringData1$listing_color <- as.numeric(as.factor(scoringData1$listing_color))
scoringData1$torque_n <- as.integer(scoringData1$torque_n)

#replace NAs with median
scoringData1$owner_count[is.na(scoringData$owner_count)] <- median(scoringData1$owner_count, na.rm = TRUE)
scoringData1$city_fuel_economy[is.na(scoringData$city_fuel_economy)] <- median(scoringData1$city_fuel_economy, na.rm = TRUE)
scoringData1$mileage[is.na(scoringData$mileage)] <- median(scoringData1$mileage, na.rm = TRUE)
scoringData1$engine_displacement[is.na(scoringData$engine_displacement)] <- median(scoringData1$engine_displacement, na.rm = TRUE)
scoringData1$highway_fuel_economy[is.na(scoringData$highway_fuel_economy)] <- median(scoringData1$highway_fuel_economy, na.rm = TRUE)
scoringData1$horsepower[is.na(scoringData$horsepower)] <- median(scoringData1$horsepower, na.rm = TRUE)
scoringData1$seller_rating[is.na(scoringData$seller_rating)] <- median(scoringData1$seller_rating, na.rm = TRUE)
scoringData1$fuel_tank_volume_gallons[is.na(scoringData$fuel_tank_volume_gallons)] <- median(scoringData1$fuel_tank_volume_gallons, na.rm = TRUE)
scoringData1$torque_n[is.na(scoringData1$torque_n)] <- median(scoringData1$torque_n, na.rm = TRUE)

scoringData2<- scoringData1%>%
  select_if(~is.numeric(.))%>%
  subset(select=-c(length_inches,width_inches,height_inches))

pred = predict(model, newdata = as.matrix(scoringData2[, features]))
submissionFile = data.frame(id = scoringData$id, price = pred)
write.csv(submissionFile, 'sample_submission.csv',row.names = F)
```

